---
title: Working with Data-Part 2-Pulling in data from Excel
author: Rob Stilson
date: '2020-09-17'
slug: working-with-data-part-2-pulling-in-data-from-excel
categories:
  - EDA
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2020-09-21T09:58:45-04:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
output:
  blogdown::html_page:
    toc: true
---

# Summary

In the last [post](http://robstilson.rbind.io/post/getting-data-from-github-part-1/) we learned how to get data from GitHub. In this post I will introduce you to getting data from an Excel file that you have on your desktop. Remember from last time we saved a file called "Additional Data for Attrition Data Set_" that included a time stamp created by R. In my case, the full file name was "Additional Data for Attrition Data Set_09_09_2020.xlsx". Let's pull it back into R using the `readxl` package from the `tidyverse`. Before we begin, remember to get and load the necessary packages.

## Get Packages

If you've read my post before, this may be repetitive, but this code chunk will load the necessary packages for the analysis. I've been using `pacman` for years and I encourage you to check it out as well. The `tryCatch` portion of the code will check your machine to see if `pacman` has already been installed. If it finds `pacman`, it will download it and if not it will move on. The `pacman::p_load` function is a wrapper for `library` and `require` and checks to see if the listed packages are installed. If they are not, it will attempt to install them from CRAN and/or any other repository in the `pacman` repository list (See *Description* under `?p_load` for additional information).

As a best practice, I also like to give a brief snippet of what each of the packages I'm loading will do in the analysis. This is helpful for anybody else who I send my code to, but also for myself if I need to come back to the code 6 months or more down the line.

```{r load packages, echo = FALSE, include= FALSE, warning = FALSE, message = FALSE}
#' <!-- ####################################################################################################### -->
#' <!-- ####################################################################################################### -->
#' <!-- ##################################LOADING PACKAGES##################################################### -->

tryCatch(require(pacman),finally=utils:::install.packages(pkgs='pacman',repos='http://cran.r-project.org'));
require(pacman)

# if the above doesn't work, use this code#
# tryCatch
# detach("package:pacman", unload = TRUE)
# install.packages("pacman", dependencies = TRUE)
# install.packages("pacman")

pacman::p_load(modeldata, # data sets useful for modeling packages
               dplyr, # a grammar of data manipulation
               tidyr, # tidy messy data
               readxl, # read Excel files
               ggplot2, # create elegant data visualisations using the grammar of graphics
               knitr, # a general-purpose package for dynamic report generation in R
               broom, # convert statistical objects into tidy tibbles
               purrr, # functional programming tools
               psych, #procedures for psychological, psychometric, and personality research
               conflicted, # an alternative conflict resolution strategy for like named function in different libraries
               janitor, # simple tools for examining and cleaning dirty data
               skimr, # compact and flexible summaries of data
               openxlsx, # read write and edit xlsx files
               RCurl, # general Network (HTTP/FTP/...) Client Interface for R
               tidyquant, # tidy quantitative financial analysis
               tidylog # logging for `dplyr` and `tidyr` functions
)

```

## Load packages

Here I'm explicitly calling the packages via `library`. Notice that `tidylog` is at the very end. This is necessary to load after `dplyr` so that all of the functions from `tidylog` work correctly. This is used in conjunction with the `conflicted` package. You can read more about it [here](https://github.com/elbersb/tidylog/blob/master/README.md).

```{r, echo = FALSE, include= FALSE, warning = FALSE, message = FALSE}

#' <!-- #Loading libraries -->

suppressPackageStartupMessages({
  library(modeldata)      
  library(dplyr)
  library(tidyr)
  library(readxl)
  library(ggplot2)
  library(knitr)
  library(broom)
  library(purrr)
  library(psych)
  library(conflicted)
  library(janitor)
  library(skimr)
  library(openxlsx)
  library(RCurl)
  library(tidyquant)
  library(tidylog, warn.conflicts = FALSE)
})

for (f in getNamespaceExports("tidylog")) {
    conflicted::conflict_prefer(f, "tidylog", quiet = TRUE)
}

```

```{r}
conflict_prefer("filter", "dplyr")
```


## Bring in the data

Now we will load the data from our desktop. You will need to change the directory to reflect where you stored the file. We will call the data `df`, which is short for data frame.

```{r}
library(readxl)
df <- read_excel("D:/Blog/Additional Data for Attrition Data Set_09_09_2020.xlsx")
```

Ok, what happened? You should now see the `df` object in your Global Environment, but where is the rest of the data? We had 8 individual tabs and it looks like it only took the first one, `accident`. What if we wanted the `performance` tab instead?

## Importing a specific worksheet from Excel

In order to do that we need to specify the sheet as follows:

```{r}
library(readxl)
performance <- read_excel("D:/Blog/Additional Data for Attrition Data Set_09_09_2020.xlsx", sheet = "performance")
```

Now we have the `performance` data in our Global Environment. We could do this 8 times for each of the sheets, but is there a better way?

## Importing multiple worksheets from Excel

This portion is from [Todd Peterson's RPubs site](https://rpubs.com/tf_peterson/readxl_import) where he shows you how to import multiple worksheets using the `lapply` function.

First, we will save our path as a variable. This is convenient so that we don't have to write out the path name each time.

```{r}
xl_data <- "D:/Blog/Additional Data for Attrition Data Set_09_09_2020.xlsx"
```

Now we can feed our `xl_data` containing the path into the `excel_sheets` function from `readxl`.

```{r}
tab_names <- excel_sheets(path = xl_data)
```

`tab_names` should appear in your Global Environment under values with the 8 tab names:

* accident
* fair_pay
* hr         
* hr_2        
* performance 
* recruitment 
* survey      
* survey_2 

We can see all of the names more easily by simply running `tab_names`.

```{r}
tab_names
```

If you wanted to remove the quotation marks, you could wrap `tab_names` in the `cat()` function.

```{r}
cat(tab_names)
```

Let's say the data you are working with has a large number of tabs. You could wrap `tab_names` in `cat()` and use a separator in order to see them all in a vertical fashion.

```{r}
cat(tab_names, sep = "\n")
```

The `"\n"` in the previous code signifies a line break.

You can also place additional characters, like a comma, before the slash to have a vertical list with a comma at the end.

```{r}
cat(tab_names, sep = ",\n")
```

Ok, back to actually importing the data. Again, we will use the `lapply` function which you can read more about [here](https://petewerner.blogspot.com/2012/12/using-apply-sapply-lapply-in-r.html) in the famous "Using apply, sapply, lapply in R" blog post.


```{r}
tab_names <- excel_sheets(path = xl_data)

list_all <- lapply(tab_names, function(x) read_excel(path = xl_data, sheet = x))

str(list_all)
```

Admittedly, I stumbled during this next step. I wanted to get all of the individual data frames back into the environment as they were when we originally pulled them from GitHub. We've got the list now with `list_all`, but how do we actually access the dfs within? [Stackoverflow](https://stackoverflow.com/questions/49359587/import-excel-workbook-with-multiple-sheets) to rescue! The linked post lays out exactly what I wanted to do and we will use that code below.

Name all of the data frames using our `tab_names` object.

```{r}
names(list_all) <- tab_names
```

Then use the until now unknown to me function of `list2env` to put them all in the Global Environment. Viola!

```{r}
list2env(list_all, .GlobalEnv)
```

